Practicle Machine Learning Writeup Assignment
========================================================

### Brief Introduction to this report.

There are devices such as *Jawbone Up, Nike FuelBand, and Fitbit* with which we can now possible to collect a large amount of data about personal activity relatively inexpensively.
In this report my goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants and predict the manner in which 6 participants did the exercise. This report also contains how I built my model, how I used cross validation and what I think the expected out of sample error is. At the end of this report there is a prediction for 20 different test cases.

Data for this report can be found in reference section.

#### Loading required library
```{r library, echo=TRUE,results='hide',warning=FALSE}
library(RANN)
library(caret)
```

#### Loading Data(Test and Train data)
```{r data, echo=TRUE, warning=FALSE}
pml.train <- read.csv("pml-training.csv")
pml.test <- read.csv("pml-testing.csv")
```
#### Data Prepration

splitting trainig dataset into train and test with a ratio of 7/3 of 10.

```{r}
set.seed(1234)
trainingIndex <- createDataPartition(pml.train$classe, list=FALSE, p=.7)
train = pml.train[trainingIndex,]
test = pml.train[-trainingIndex,]
```
Filter the numeric features and outcome. 
```{r}
num_idx = which(lapply(train,class) %in% c('numeric')  )
```
Fixing the  missing values. This includes the imputation method using K-nearest neighbour model.

```{r}
preModel <- preProcess(train[,num_idx], method=c('knnImpute'))
ptraining <- cbind(train$classe, predict(preModel, train[,num_idx]))
ptesting <- cbind(test$classe, predict(preModel, test[,num_idx]))
```
Rename first Label to classe

```{r}
names(ptraining)[1] <- 'classe'
names(ptesting)[1] <- 'classe'
```

Make test set for submission

```{r}
prtesting <- predict(preModel, pml.test[,num_idx])
```
### Modeling with Random Forest Model

```{r, warning=FALSE}
library(randomForest)
rf_model  <- randomForest(classe ~ ., ptraining)
plot(rf_model)
```
### Important Variables for above applied model
in this list variables with higher value are more important for above model

```{r, warning=FALSE}
varImp(rf_model)
``` 

### In-sample accuracy
```{r}
training_pred <- predict(rf_model, ptraining) 
print(confusionMatrix(training_pred, ptraining$classe))
```
The in-sample accuracy is 100%.

### Out-of-sample accuracy
```{r}
testing_pred <- predict(rf_model, ptesting) 
```

Confusion Matrix for train dataset: 
```{r}
print(confusionMatrix(testing_pred, ptesting$classe))
```

The cross validation accuracy is  98.6%, which should be sufficient for predicting the 20 test observations.

### Test Set Prediction Results

Applying model to the test data.
```{r}
answers <- predict(rf_model, prtesting) 
answers
```

Writing outputs to 20 files.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```

## References
[1] http://groupware.les.inf.puc-rio.br/har#weight_lifting_exercises  
[2] https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv  
[3] https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv